### 2021-05-26 LaTeX-ACM/sections/paper.tex
edits + resolving merge conflicts

In designing \cm within Ethereum's gas model, performance is the main bottleneck. For a call market, closing the market and processing all the orders are the most time-consuming steps. The most critical design decision is the data structure for holding orders. While data structures are well studied for many languages, Solidity/EVM has its own unique aspects (\eg gas refunds, a relatively cheap mapping data structure, only partial support for object-oriented programming) that create difficulties in assessing which will perform best without actually deploying and evaluating each variant.

When closing a call market, the orders are examined in order: highest to lowest price for bids, and lowest to highest price for asks. In most circumstances, the market closing algorithm does not have to consider any deeper bids/asks from the list when choosing whether the current best bid and ask can be fulfilled. The only exceptions are in the case of a tie on price or a canceled order, both of which we return to later. For this reason, the ideal data structure for storing bids/asks is a \textit{priority queue} (see Table~\ref{tab:PQ_API}) where each order's priority is its price. Specifically, we use two PQs---one for bids where the highest price is the highest priority, and one for asks where the lowest price is the highest priority.

@@

\item \textbf{Linked List.} In this variant, we insert a new element into its correct position (based on its price) when running enqueue. The PQ itself stores elements in a linked list (enabling us to efficiently insert a new element between two existing elements). Solidity is described as object-oriented but the equivalent of an object is an entire smart contract. Therefore an object-oriented linked list must either (i) create each node in the list as a struct---but this is not possible as Solidity does not support recursive structs---or (ii) make every node in the list its own contract. The latter option seems wasteful and unusual, but we try it out anyways. Thus each node is its own contract and contains the order data and a pointer to the address of the next contract in the list.

\item \textbf{Linked List with Mapping.} Finally, we try a variant of a linked list using a Solidity mapping. The value of the mapping is a struct with order data and the key of the next (and previous) node in the list. The contract stores the key of the first node (head) and last node (tail) in the list.

@@

\paragraph{Enqueue Performance.} We implemented, deployed, and tested each PQ using Truffle and Ganache. We tried a variety of tests (including testing the full-call market with each variant) with consistent results in performance. A simple test to showcase the performance profile is shown in Figure~\ref{fig:random_insertion}. We simply enqueue 50 integers chosen at random from a fixed interval in each PQ variant. The bigger the PQ gets, the longer enqueue takes---a linear increase for the linked list variants, and logarithmic for the heap variants.

@@

\paragraph{Discussion.} Based on enqueuing, the heap variants are the cheapest in terms of gas, while based on dequeuing, the link list variants are the cheapest. This is in accordance with the theoretical worst-case time complexity for each. However, (i) the linked list variants are materially cheaper than the heap variants at dequeuing, and (ii) dequeuing in a call market must be done as a batch, whereas enqueuing is paid for one at a time by the trader submitting the order, and (iii) Ethereum will not permit more than hundreds of orders so the asymptotic behavior is not a significant factor. For these reasons, we suggest using a linked list variant for this specific application. As it can be seen in Figure~\ref{fig:random_insertion}, the associated cost for inserting elements into a linked list PQ is significantly greater than the linked list with mapping, as each insertion causes the creation of a new DApp. Accordingly, we choose to implement the call market with the linked list with mapping. Overall this PQ balances a moderate gas cost for insertion (\ie order submission) with one for removal (\ie matching the orders).

@@

\paragraph{Gas Refunds.} In order to reconstruct the current state of Ethereum's EVM, a node must obtain a copy of every variable change since the genesis block (or a more recent `checkpoint' that is universally agreed to). For this reason, stored variables persist for a long time and, at first glance, it seems pointless to free up variable storage (and unclear what `free up' even means). Once the current state of the EVM is established by a node, it can forget about every historical variable changes and only concern itself with the variables that have non-zero value (as a byte string for non-integers) in the current state (uninitialized variables in Ethereum have the value 0 by default). Therefore, freeing up variables will reduce the amount of state Ethereum nodes need to maintain going forward.

@@

\item \texttt{SELF-DESTRUCT}. This operation destroys the contract that calls it and refunds its balance (if any) to a designated receiver address. The  \texttt{SELFDESTRUCT} operation does not remove the initial byte code of the contract from the chain. It always refunds 24,000 gas. For example, if contract A stores a single non-zero integer and contract B stores 100 non-zero integers, the \texttt{SELFDESTRUCT} refund for both is the same (24,000 gas).

@@

Dequeuing in a linked list with mapping can be implemented in two ways. The simplest approach is to process a node, update the head pointer, and leave the `removed' node's data behind in the mapping untouched (where it will never be referenced again). Alternatively, we can call \texttt{DELETE} on each mapping entry once we are done processing a node in the PQ. As it can be seen in the last two rows of Table~\ref{tab:cleaning}, leaving the data on chain is cheaper than cleaning it up.

@@

As expected, the numbers closely track the performance of the PQ itself suggesting the PQ is indeed the main influence on performance. In Ethereum today, call markets appear to be limited to processing about a hundred orders per transaction. If markets open on every other block and the call market could monopolize an entire block to close, a few hundred orders per minute (worst-case) can be processed. The main takeaway is that the transparency, front-running resistance, and low barrier to entry of Ethereum come with an enormous cost (\ie an institutional exchange like NASDAQ can process 100K trades per second). That said, many exchanges trade the same assets under different trading rules (\ie market fragmentation) because traders have different preferences. \cm can work today in some circumstances like very low liquidity tokens, or markets with high volumes and a small number of traders (\eg liquidation auctions).

@@

Beyond the cleaning up issues with priority queues in Section~\ref{sec:gasrefund}, \cm also uses mappings with each market. Traders preload their account with tokens to be traded (which comply with a common token standard called ERC20) and/or ETH. \cm tracks what they are owed using a mapping called \texttt{totalBalance} and allows traders to withdraw their tokens at any time. However if a trader submits an order (\ie ask for their tokens), the tokens are committed and not available for withdrawal until the market closes (after which, the balances are updated for each trade that is executed). Committed tokens are also tracked in a mapping called \texttt{unavailableBalance}. Sellers can request a token withdrawal up to their total balance subtracted by their unavailable balance.

As the DApp runs \texttt{closeMarket()}, it starts matching the best bids to the best asks. As orders execute, \texttt{totalBalance} and \texttt{unavailableBalance} are updated. At a certain point, the bids and asks will stop matching in price. At this point, every order left in the order book cannot execute (because the priority queue sorts orders by price, and so orders deeper in the queue have worst prices than the order at the head of the queue). Therefore all remaining entries in \texttt{unavailableBalance} can be cleared.

@@

\item \textbf{Store the Mapping in a Separate DApp.} We could wrap the mapping inside its own DApp and when we are done with the mapping, we can run \texttt{SELFDESTRUCT} on the contract. This refunds us 24,000 gas which is less than the cost of deploying the extra contract. Additionally, every call to the mapping is more expensive because (1) it is an external function call, and (2) the calls need access control to ensure only the market contract can write to it (if a mapping is a local variable, you get private access for free).

@@

Call markets are heralded for fair price discovery. This is why many exchanges use a call market at the end of the day to determine the closing price of an asset, which is an important price both optically (it is well published) and operationally (many derivatives settle based on the closing price). We purposely do not compute a `market clearing price' with \cm because miners can easily manipulate the price (\ie include a single wash trade at the price they want fixed), although they forgo profit for doing so. This is not merely hypothetical---Uniswap (the prominent quote-drive, on-chain exchange) prices have been manipulated to exploit other DeFi applications relying on them. Countermeasures to protect Uniswap price integrity could also apply to \cm: (1) taking a rolling median of prices over time, and (2) using it alongside other sources for the same price and forming a consensus. While \cm does not emit a market clearing price, it can be computed by a web application examining the order book at market close.

@@

In \cm, both the tokens and ETH that a trader want to potentially use in the order book are pre-loaded into the contract. Consider Alice, who holds a token and decides she wants to trade it for ETH. In this model, she must first transfer the tokens to the contract and then submit an ask order. If she does this within the same block, there is a chance that a miner will execute the ask before the transfer and the ask will revert. If she waits for confirmation, this introduces a delay. This delay seems reasonable but we point out a few options it could be addressed:

@@

Arbitrum is an Ethereum Layer 2 scaling solution that operates as an optimistic roll-up~\cite{kalodner2018arbitrum}. Arbitrum runs entire DApps inside its environment, which acts like a container and requires special bridge contacts to interact with `normal' (or L1) Ethereum DApps. To deploy a DApp on Arbitrum, or to execute a function on an existing Arbitrum DApp, the transaction is sent to an L1 Arbitrum contract called an inbox. The transaction is not executed, it is only recorded (as calldata) by the inbox contract. Like any other transactions on Ethereum, miners determine the order of the transactions sent to the inbox. A set of validators (open to anyone) are watching the inbox contract for new transactions. Once an inbox transaction is finalized in an Ethereum block, validators will execute the transaction and asserts the result of the execution to other validators on a sidechain called ArbOS. Once the dispute challenge period is over, the new state of ArbOS will be inserted into Ethereum. As the inbox contract maintains all Arbitrum transactions, anyone can recompute the entire current state of the ArbOS and file a dispute if executions are not performed correctly. Disputes are adjudicated by Ethereum itself and require essentially constant gas costs, regardless of the size of the gas costs of the transaction being disputed.

@@

\textit{Testing Platforms.} Arbitrum currently operates a test network with an inbox contract on the Ethereum Kovan testnet. To experiment with Arbitrum, we implemented a variant of \cm and run the same experiments that we previously performed on Layer 1 (Ethereum). To call functions on \cm, traders can (i) send transactions directly to the inbox contract \href{https://kovan.etherscan.io/address/0x76bF1345224fE606E2aB38B8E52B83512328A9DF}, or (ii) use a relay server (called a \textit{Sequencer}) provided by the Arbitrum. The sequencer will group, order,  and send all pending transactions together as a single Kovan transaction to the inbox (and pay the gas).

\cm implements an exchange between ETH and tokens. To use Layer 1 assets on Layer 2, Arbitrum provides two bridge contracts: one for ETH, and one for tokens compliant with ERC-20 or other standard interfaces. Prior to trading, a trader uses the bridge to move their ETH to the same address on Layer 2 and then move the ETH to \cm with a Layer 2 transaction. From the user's perspective, tokens work similarly although the backend details are more complex. Finally, withdrawals work essentially same way in reverse, but are only final on Layer 1 after a dispute challenge period (currently 1 hour).\footnote{Layer 1 users might accept assets before they are finalized if they determine their eventual emergence on Layer 1 is indisputable (\textit{eventual finality}).}

@@

In our \cm variant on Arbitrum, the token issuer does all the computation (both enqueuing and dequeuing). Thus we switch the priority queue to use a heap with dynamic array, which balances the expense of both operations (instead of optimizing for dequeuing in \texttt{closeMarket()}). On Arbitrum, heaps are 32\% more efficient than non-Heaps for submitting orders and 29\% less efficient for closing. Recall that without a roll-up, such a priority queue can only match 38 pairs at a cost of 5,372,679 gas. With Arbitrum roll-ups, 38 pairs costs only 1,632 in L1 gas (a 99.9\% savings). This is effectively the cost of recording the \texttt{closeMarket()} function call in the inbox, which is 103 bytes of calldata. As the number of pairs increases, L1 gas cost is constant. For traders submitting an order, the cost is reduced from  207,932 to 2,656.

@@

While the runtime of a transaction is no longer an issue for Ethereum, it is still an issue for the validators. Like Ethereum miners, validators do measure the amount of gas (called ArbGas in Arbitrum) a transaction takes although Arbitrum emulates EVM code with its own different OpCodes. Thus, ArbGas with a roll-up and L1 gas without a roll-up are not directly comparable but both increase and decrease proportionately with the complexity of running the function. Closing a market with 38 pairs costs 46,705,693 ArbGas. Until Arbitrum deploys on mainnet, the price of ArbGas in Gwei (or USD) is unknown but it is thought to be relatively very cheap.

>>>>

In designing \cm within Ethereum's gas model, performance is the main bottleneck. For a call market, closing the market and processing all the orders are the most time-consuming steps. The most critical design decision is the data structure for holding orders. While data structures are well-studied for many languages, Solidity/EVM has its own unique aspects (\eg gas refunds, a relatively cheap mapping data structure, only partial support for object-oriented programming) that create difficulties in assessing which will perform best without actually deploying and evaluating each variant.

When closing a call market, the orders are examined in order: highest to lowest price for bids, and lowest to highest price for asks. In most circumstances, the market closing algorithm does not have to consider any deeper bids/asks from the list when choosing whether the current best bid and ask can be fulfilled. The only exceptions are in the case of a tie on price or a canceled order, both of which we return to later. For this reason, the ideal data structure for storing bids/asks is a \textit{priority queue} (see Table~\ref{tab:PQ_API}) where each order's priority is its price. Specifically, we use two PQs---one for bids, where the highest price is the highest priority, and one for asks, where the lowest price is the highest priority.

@@

\item \textbf{Linked List.} In this variant, we insert a new element into its correct position (based on its price) when running enqueue. The PQ itself stores elements in a linked list (enabling us to efficiently insert a new element between two existing elements). Solidity is described as object-oriented but the equivalent of an object is an entire smart contract. Therefore, an object-oriented linked list must either (i) create each node in the list as a struct---but this is not possible as Solidity does not support recursive structs---or (ii) make every node in the list its own contract. \textblue{The latter option seems wasteful and unusual, but we try it out anyways.} Thus, each node is its own contract and contains the incoming order's data and a pointer to the address of the next contract in the list.

\item \textbf{Linked List with Mapping.} Finally, we try a variant of a linked list using a Solidity mapping. The value of the mapping is a struct with the incoming order's data and the key of the next (and previous) node in the list. The contract stores the key of the first node (head) and last node (tail) in the list.

@@

\paragraph{Enqueue Performance.} We implemented, deployed, and tested each PQ using the Truffle development framework~\cite{TruffleO71:online} and Ganache~\cite{GanacheT25:online}. We tried a variety of tests (including testing the full call market with each variant) with consistent results in performance. A simple test to showcase the performance profile is shown in Figure~\ref{fig:random_insertion}. We enqueue 50 integers chosen at random from a fixed interval in each PQ variant. The bigger the PQ gets, the longer enqueue takes---a linear increase for the linked list variants, and logarithmic for the heap variants.

@@

\paragraph{Discussion.} Based on enqueuing, the heap variants are the cheapest in terms of gas, while based on dequeuing, the link list variants are the cheapest. This is in accordance with the theoretical worst-case time complexity for each. However, (i) the linked list variants are materially cheaper than the heap variants at dequeuing, and (ii) dequeuing in a call market must be done as a batch, whereas enqueuing is paid for one at a time by the trader submitting the order, and (iii) Ethereum will not permit more than hundreds of orders so the asymptotic behavior is not a significant factor. For these reasons, we suggest using a linked list variant for this specific application. As it can be seen in Figure~\ref{fig:random_insertion}, the associated cost for inserting elements into a linked list PQ is significantly greater than the linked list with mapping, as each insertion causes the creation of a new DApp. Accordingly, we choose to implement the call market with the linked list with mapping. Overall this PQ balances a moderate gas cost for insertion (\ie order submission) with one for removal (\ie closing the market and matching the orders).

@@

\paragraph{Gas Refunds.} In order to reconstruct the current state of Ethereum's EVM, a node must obtain a copy of every variable change since the genesis block (or a more recent `checkpoint' that is universally agreed to). For this reason, stored variables persist for a long time and, at first glance, it seems pointless to free up variable storage (and unclear what `free up' even means). Once the current state of the EVM is established by a node, it can forget about every historical variable changes and only concern itself with the variables that have non-zero value (as a byte string for non-integers) in the current state (uninitialized variables in Ethereum have the value 0 by default).\textblue{ Therefore, freeing up variables will reduce the amount of state Ethereum nodes need to maintain going forward.}

@@

\item \texttt{SELFDESTRUCT}. This operation destroys the contract that calls it and refunds its balance (if any) to a designated receiver address. The  \texttt{SELFDESTRUCT} operation does not remove the initial byte code of the contract from the chain. It always refunds 24,000 gas. For example, if contract A stores a single non-zero integer and contract B stores 100 non-zero integers, the \texttt{SELFDESTRUCT} refund for both is the same (24,000 gas).

@@

Dequeuing in a linked list with mapping can be implemented in two ways. The simplest approach is to process a node, update the head pointer, and leave the `removed' node's data behind in the mapping untouched (where it will never be referenced again). Alternatively, we can call \texttt{DELETE} on each mapping entry once we finish processing a node in the PQ. As it can be seen in the last two rows of Table~\ref{tab:cleaning}, leaving the data on chain is cheaper than cleaning it up.

@@

As expected, the numbers closely track the performance of the PQ itself, suggesting the PQ is indeed the main influence on performance. In Ethereum today, call markets appear to be limited to processing about a hundred orders per transaction. If markets open on every other block and the call market could monopolize an entire block to close, a few hundred orders per minute (worst-case) can be processed. The main takeaway is that the transparency, front-running resistance, and low barrier to entry of Ethereum come with an enormous cost (\ie an institutional exchange like NASDAQ can process 100K trades per second). That said, many exchanges trade the same assets under different trading rules (\ie market fragmentation) because traders have different preferences. \cm can work today in some circumstances like very low liquidity tokens, or markets with high volumes and a small number of traders (\eg liquidation auctions).

@@

\textblue{Beyond the cleaning up issues with priority queues in Section~\ref{sec:gasrefund}, \cm also uses mappings with each market.} Traders preload their account with tokens to be traded (which comply with a common token standard called ERC20) and/or ETH. \cm tracks what they are owed using a mapping called \texttt{totalBalance} and allows traders to withdraw their tokens at any time. However if a trader submits an order (\ie ask for their tokens), the tokens are committed and not available for withdrawal until the market closes (after which, the balances are updated for each trade that is executed). Committed tokens are also tracked in a mapping called \texttt{unavailableBalance}. Sellers can request a token withdrawal up to their total balance subtracted by their unavailable balance.

\textblue{As the DApp runs \texttt{closeMarket()}}, it starts matching the best bids to the best asks. As orders execute, \texttt{totalBalance} and \texttt{unavailableBalance} are updated. At a certain point, the bids and asks will stop matching in price. At this point, every order left in the order book cannot execute (the priority queue sorts orders by price, so orders deeper in the queue have worst prices than the order at the head of the queue). Therefore all remaining entries in \texttt{unavailableBalance} can be cleared.

@@

\item \textbf{Store the Mapping in a Separate DApp.} We could wrap the mapping inside its own DApp and run \texttt{SELFDESTRUCT} on the contract once we finish processing the mapping. This refunds us 24,000 gas which is less than the cost of deploying the extra contract. Additionally, every call to the mapping is more expensive because (1) it is an external function call, and (2) the calls need access control to ensure only the market contract can write to it (if a mapping is a local variable, you get private access for free).

@@

Call markets are heralded for fair price discovery. This is why many exchanges use a call market at the end of the day to determine the closing price of an asset, which is an important price both optically (it is well-published) and operationally (many derivatives settle based on the closing price). We purposely do not compute a `market clearing price' with \cm because miners can easily manipulate the price (\ie include a single wash trade at the price they want fixed), although they forgo profit for doing so. This is not merely hypothetical---Uniswap (the prominent quote-drive, on-chain exchange) prices have been manipulated to exploit other DeFi applications relying on them. Countermeasures to protect Uniswap price integrity could also apply to \cm: (1) taking a rolling median of prices over time, and (2) using it alongside other sources for the same price and forming a consensus. While \cm does not emit a market clearing price, it can be computed by a web application examining the order book at market close.

@@

In \cm, both the tokens and ETH that a trader wants to potentially use in the order book are pre-loaded into the contract. Consider Alice, who holds a token and decides she wants to trade it for ETH. In this model, she must first transfer the tokens to the contract and then submit an ask order. If she does this within the same block, there is a chance that a miner will execute the ask before the transfer and the ask will revert. If she waits for confirmation, this introduces a delay. This delay seems reasonable but we point out a few options it could be addressed:

@@

Arbitrum is an Ethereum Layer 2 scaling solution that operates as an optimistic roll-up~\cite{kalodner2018arbitrum}. Arbitrum runs entire DApps inside its environment, which acts like a container and requires special bridge contacts to interact with `normal' (or Layer 1) Ethereum DApps. To deploy a DApp on Arbitrum, or to execute a function on an existing Arbitrum DApp, the transaction is sent to a Layer 1 Arbitrum contract called Inbox. The transaction is not executed, it is only recorded (as calldata) by the Inbox contract. Like any other transactions on Ethereum, miners determine the order of the transactions sent to this contract. A set of validators (open to anyone) are watching the Inbox contract for new transactions. Once an inbox transaction is finalized in an Ethereum block, validators will execute the transaction and asserts the result of the execution to other validators on a sidechain called ArbOS. Once the dispute challenge period is over, the new state of ArbOS will be inserted into Ethereum. As the Inbox contract maintains all Arbitrum transactions, anyone can recompute the entire current state of the ArbOS and file a dispute if executions are not performed correctly. Disputes are adjudicated by Ethereum itself and require essentially constant gas costs, regardless of the size of the gas costs of the transaction being disputed.

@@

\textit{Testing Platforms.} Arbitrum currently operates a test network with an Inbox contract on the Ethereum Kovan testnet. To experiment with Arbitrum, we implemented a variant of \cm and ran the same experiments that we previously performed on Layer 1 (Ethereum). To call functions on \cm, traders can (i) send transactions directly to the Inbox contract \href{https://kovan.etherscan.io/address/0x76bF1345224fE606E2aB38B8E52B83512328A9DF}, or (ii) use a relay server (called a \textit{Sequencer}) provided by the Arbitrum. The sequencer will group, order,  and send all pending transactions together as a single Kovan transaction to the Inbox (and pay the gas).

Recall that \cm implements an exchange between ETH and tokens. To use Layer 1 assets on Layer 2, Arbitrum provides two bridge contracts: one for ETH, and one for tokens compliant with ERC-20 or other standard interfaces. Prior to trading, a trader uses the bridge to move their ETH to the same address on Layer 2 and then move the ETH to \cm with a Layer 2 transaction. From the user's perspective, tokens work similarly although the backend details are more complex. Finally, withdrawals work essentially same way in reverse, but are only final on Layer 1 after a dispute challenge period (currently 1 hour).\footnote{Layer 1 users might accept assets before they are finalized if they determine their eventual emergence on Layer 1 is indisputable (\textit{eventual finality}).}

@@

In our \cm variant on Arbitrum, the token issuer does all the computation (both enqueuing and dequeuing). Thus, we switch the priority queue to use a heap with dynamic array, which balances the expense of both operations (instead of optimizing for dequeuing in \texttt{closeMarket()}). Our experiments show that on Arbitrum, heaps are 32\% more efficient than non-Heaps for submitting orders and 29\% less efficient for closing. Recall that without a roll-up, such a priority queue can only match 38 pairs at a cost of 5,372,679 gas. With Arbitrum roll-ups, 38 pairs cost only 1,632 in Layer 1 gas (a 99.9\% savings). This is effectively the cost of recording the \texttt{closeMarket()} function call in the Inbox, which is 103 bytes of calldata. As the number of pairs increases, Layer 1 gas cost is constant. For traders submitting an order, the cost is reduced from  207,932 to 2,656.

@@

While the runtime of a transaction is no longer an issue for Ethereum, it is still an issue for the validators. Similar to Ethereum miners, validators do measure the amount of gas (called ArbGas in Arbitrum) a transaction takes, although Arbitrum emulates EVM code with its own different opcodes. Thus, ArbGas with a roll-up and Layer 1 gas without a roll-up are not directly comparable but both increase and decrease proportionately with the complexity of running the function. Closing a market with 38 pairs costs 46,705,693 ArbGas. Until Arbitrum deploys on mainnet, the price of ArbGas in Gwei (or USD) is unknown but it is thought to be relatively very cheap.
